---
title: 'Introduction to ramlegacy'
author: "Carl Boettiger & Kshitiz Gupta" 
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: |
  %\VignetteIndexEntry{ramlegacy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## RAM Legacy Stock Assessment Database
From the database's [website](www.ramlegacy.org):

> The RAM Legacy Stock Assessment Database is a compilation of stock assessment results for commercially exploited marine populations from around the world. It is inspired by Dr. Ransom A. Myers' original [stock-recruitment database](http://www.mathstat.dal.ca/~myers/welcome.html), which is no longer being updated."
Different versions of the database are available in compressed (zipped) form on the [website](www.ramlegacy.org) in two formats: **Access** (Microsoft Office Access 2007 file format)  and **Excel** (Microsoft Excel Open XML Format Spreadsheet). As of March 2019, the available versions are 1.0, 2.0, 2.5, 3.0, 4.3, and 4.40, 4.41, 4.44.

## About the package
The goal of `ramlegacy` is to cache the specified version/versions of the RAM Legacy Stock Assessment Excel Database and this way save time and effort spent in re-downloading different
versions of the database as part of any future analysis involving the data. The package also supports reading in any specified dataframe(s) of the database through a function called `load_ramlegacy()`. The user can always view the location to which the database is cached and read from by running `ram_dir()` and specifying its version argument. This vignette will outline a few more details about these functions and the package that will hopefully make `ramlegacy` easy to use.

## Package Installation
You can install the development version from [GitHub](https://github.com/kshtzgupta1/ramlegacy) with the `install_github` function available in `devtools` package:

```{r, eval = FALSE, echo = TRUE}
install.packages("devtools")
library(devtools)
install_github("kshtzgupta1/ramlegacy")
```

To ensure that the vignette is installed along with the package make sure to remove `--no-build-vignettes` from the `build_opts` in `install_github`

## download_ramlegacy
You can download any available version of the RAM Legacy Stock Assessment Excel Database through `download_ramlegacy()` and specifying the version argument:

```{r, download_ramlegacy_example1, eval=FALSE, echo = T}
# download version 1.0
download_ramlegacy(version = "1.0")
# download version 2.0
download_ramlegacy(version = "2.0")
# download version 2.5
download_ramlegacy(version = "2.5")
# download version 3.0
download_ramlegacy(version = "3.0")
# download version 4.3
download_ramlegacy(version = "4.3")
```

If version is not specified then `download_ramlegacy` will default to downloading the
latest version of the database:

```{r, download_ramlegacy_example2, echo = T, eval = F}
# downloads latest version (currently 4.44)
download_ramlegacy()
```
If you want to download multiple versions please download them one at a time as passing them all at once will throw an error.

To enable caching behavior, before downloading a the specified version `download_ramlegacy` checks whether that version has already been downloaded. If it is then `download_ramlegacy` gives the user the option to either download again and overwrite the existing version or not download and just exit the function. This behavior is modified if `overwrite`, a function argument to `download_ramlegacy`, is set to true. Then even if the version has already been downloaded calling `download_ramlegacy` will download it again overwriting the already present version. After downloading the database download_ramlegacy caches it by converting it to an RDS object which can then be read in through calling `load_ramlegacy()`. 

`download_ramlegacy` also has three other arguments: `ram_path`, `ram_url`, and `quiet`. By default `ram_path` has been set to a location provided by the rappdirs package and can be viewed by calling `ram_dir(vers = "version")`. Although this is not the recommended approach download_ramlegacy supports downloading to a user-specified path. `ram_url` is fixed to the zenodo URL of the database. Please do not **pass in any other url** to ram_url. By default, `download_ramlegacy` after it is called keeps the user updated of the download progress through status messages. Setting `quiet` to TRUE enables the user suppress all status messages.
  
## load_ramlegacy
`load_ramlegacy` can be used to obtain a list of particular dataframes and passed in to the `tables` argument. To get a list of all the dataframes within a specific version of the database set tables = NULLired version of the database. Make sure that the version you want to load was downloaded and is already present locally otherwise `load_ramlegacy` will throw an error. 

```{r, load_ramlegacy_example1, echo = T, eval = F}
# returns a list containing area and bioparams tables from version 4.44 database
load_ramlegacy(version = "4.44", tables = c("area", "bioparams"))
```

In case in you want to load multiple versions please load them one at a time as passing them all at once will throw an error. Similar, to `download_ramlegacy`, `load_ramlegacy` has a `ram_path` argument which is set to local directory where the specified version of the RAM Legacy Stock Excel Assessment Database was downloaded. You can view that path through `ram_dir(vers = "version")`. Although this is not the recommended approach load_ramlegacy supports reading in the database's dataframes from a user-specified path provided that the database is present at the specified path as an rds file.
  
## ram_dir
To see the path where the specified version of database was downloaded on the user's computer, the user can call `ram_dir`. Note that this is also the location from which `load_ramlegacy` reads in the database. 

For example, to see where version 4.3 of the database was downloaded you can call:

```{r, ram_dir_example1, echo = T, eval = F}
# view the path where version 4.3 of the database was cached
ram_dir(vers = "4.3")
```
